A community-maintained standard library of population genetic models

### 2019-12-20 popSimManu.tex
edits until the Methods, including ones from #112
finished edits

fixed supp fig referencing, closes #109

number of samples; closes #128

what is top; closes #129

responses to edits; closes #124

url fix

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
\newcommand{\stopsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{\arabic{figure}}%
     }

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but specifying such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and the possibility for error.
Population genetics, as a field, also lacks standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, \stdpopsim, that attempts to rectify this situation.
\Stdpopsim is a community-driven open source project, which provides a standard
catalog of published simulation models from a wide range of organisms
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature, community-based quality control pipeline.
We share some examples demonstrating how \stdpopsim
can be used to systematically compare demographic inference methods,
and we encourage an even broader
community of developers to join in the effort to standardize
population genetic methods.}
\date{}

\keywords{Population genetics, Simulation, Inference}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{Boyko:2008cr,kim2017inference,tataru2017inference,Fortier703918,Huang2019,Vecchyo770966},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247,Barroso2019},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming ever more important.
%CK: this first paragraph is super long - might be helpful to break up

Assessing the accuracy of inference methods for population genetics is
challenging in large part because the ``ground-truth'' in question
generally comes not from direct empirical observations, as the relevant
historical processes can rarely be observed, but instead from simulations.
Population genetic simulations are therefore critically important to the
field, yet there has been no systematic attempt to establish community
standards or best practices for executing them. Instead, the general modus
operandi to date has been for individual groups to validate their own
methods using bespoke simulations. 
Often these studies focus more on showcasing a novel method than on
rigorously comparing it with competing methods.
Moreover, this approach results in a great deal of duplicated effort across research
groups, and contributes to decreased reproducibility and transparency.
It is also a barrier to entry to the field, because new researchers can
struggle with 
the many steps involved
in implementing a state-of-the-art population genetics simulation,
including identifying appropriate demographic models from the literature, 
translating them into input for a simulator, obtaining appropriate genetic
maps, and choosing appropriate values for key population genetic parameters.

A related issue is that it has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect population
genetic inferences. Yet there
are clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, some methods are likely to
perform better than others under certain scenarios, but 
researchers lack principled guidelines for selecting the best method for addressing 
their particular questions. The need for empirical
guidance will only increase as researchers 
seek to apply population genetic methods to a growing collection of non-model taxa.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains three organisms: humans,
\emph{Drosophila~melanogaster}, and \emph{Arabidopsis~thaliana}. For each
organism, the catalog contains details on the physical organization (e.g., chromosome structure)
of its genome, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.
In this way, \stdpopsim will enable rigorous evaluation of population genetic software,
and contribute to increased reliability of population genetic inferences.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/Figure1.pdf}
    \caption{\textbf{Structure of \stdpopsim}. \textbf{(A)} The
hierarchical organization of the \stdpopsim catalog contains all model simulation information
within individual species (expanded information shown here for \emph{H. sapiens} only). 
Each species is associated with a representation of the physical genome, and one or more genetic maps and demographic models.
Dotted lines indicate that only a subset of these categories is shown. \textbf{(B)} Example code to specify
and simulate models using \stdpopsim via the API or \textbf{(C)} through a command line interface.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library.}
The first contribution of the PopSim consortium is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans,
\emph{D.~melanogaster}, \emph{A.~thaliana}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about each species' genome, including information about chromosome
lengths, average mutation rates, and generation times. We also provide access to detailed
empirical information such as genetic maps, which model observed
heterogeneity in recombination rate along chromosomes. As such maps are often large,
we do not distribute them directly with the software, but make them available
for download in a standard format. When a simulation using such a map is
requested by the user, \stdpopsim will transparently download the map
data into a local cache (\texttt{$\sim$/.local/cache/stdpopsim} by default
on Unix platforms), where it can be quickly retrieved for subsequent
simulations.
In the initial version of \stdpopsim we support
the HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps in humans;
the \cite{salome2011recombination} map in \emph{A.~thaliana};
and the \cite{comeron2012many} map in \emph{D.~melanogaster}.
The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic variation. (See the Methods for a description of the community
development and quality-control process for these models.)

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modeled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim} as
an additional backend.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work, and encourage authors
of simulation models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number and the long-term stable identifiers
for population models and genetic maps,
this citation information will result in well-documented and reproducible
simulation workflows. The individual tree sequence files produced by
\stdpopsim also contain complete provenance information including the command
line arguments, operating system environment and versions of key libraries
used.

@@

The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}, 
and we have models in development for \emph{Pongo abelii} and \emph{Escherichia coli}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

\renewcommand{\arraystretch}{1.2}
\begin{table}[t]
% Make sure the table is centered even though it's too wide
\makebox[\textwidth][c]{
    \begin{footnotesize}
    \input{model_table.tex}
    \end{footnotesize}
}
\caption{\label{tab:catalog}
Initial set of demographic models in the Catalog and simple benchmarks.
For each model we report the CPU time, maximum memory usage and the
size of the output \tskit file. In each case we simulate 100 samples
drawn from the first population, for the shortest chromosome of that species
and a constant chromosome-specific recombination rate.
The times reported are for a single replicate run on an Intel i5-7600 CPU.
These are not intended to be rigorous benchmarks, and the computing resources
required will vary widely depending on sample sizes, chromosome length,
recombination rates and other factors.
}
\end{table}

Currently, {\em Homo sapiens} has the largest number of population models in
\stdpopsim (see Table \ref{tab:catalog}).  
These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth; Africa\_1T12), the three-population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations (OutOfAfrica\_3G09), the \cite{tennessen2012evolution} two-population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth (OutOfAfrica\_2T12), the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components (AmericanAdmixture\_4B11), a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture (OutOfAfricaArchaicAdmixture\_5R19),
a complex model of ancient Eurasian admixture from \cite{kamm2019efficiently} (AncientEurasia\_9K19),
and a synthetic model of oscillating population size from \cite{schiffels2014inferring} (Zigzag\_1S14).
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{D. melanogaster}, three from \emph{A. thaliana}, and have under development
model for \emph{P. abelii} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample (African3Epoch\_1S16), as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring}, which jointly models African
and European populations (OutOfAfrica\_2L06). For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes (SouthMiddleAtlas\_1D17). We have also implemented a two-epoch and a three-epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene} (African2Epoch\_1H18 and African3Epoch\_1H18). 
%For \emph{P. pygmaeus} we have implemented the two
%species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
%and \emph{P. abelii}. 
%For \emph{E. coli} we have a simple, large constant population size model implemented.
Beyond organism specific models \stdpopsim also includes a generic piecewise-constant-size model and
an isolation-with-migration (IM) model, both of which can be used with any genome and genetic-map combination.

We note that we have put a good deal of effort into coming up with a standardized naming
conventions for species, genetic maps, and demographic models that will enable long term stability
of unique identifiers used throughout \stdpopsim. This is an essential feature to guarantee reproducibility.
\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparisons of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community
(see Discussion).

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To reproducibly evaluate and compare the performance of inference methods, we developed
\texttt{snakemake} \citep{koster2012snakemake} workflows,
available from \url{https://github.com/popsim-consortium/analysis}.
These workflows allow efficient computing in multicore or cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $n$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairway plot}, and
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the OutOfAfricaArchaicAdmixture\_5R19 model and using the
HapMapII\_GRCh37 genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model (see Appendix).}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale}, we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under OutOfAfricaArchaicAdmixture\_5R19,
a model of human migration out of Africa that includes archaic admixture 
\citep{ragsdale2019models}, along with an empirical genetic map. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC).
There is no single ``true'' reference for effective population size
because of model misspecification---the inference methods are fitting a single population model
to data simulated from multiple populations.  
However, many methods work by matching coalescence time distributions,
and a single-population model with varying population size can match any coalescence time distribution
(in which case coalescence rate is the inverse of the effective size).
For this reason, we used as our ``ground-truth'' (solid black lines) 
not historical census sizes,
but rather inverse coalescence rates calculated analytically in \texttt{msprime} (see Appendix).
Blue lines show estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
the methods are generally accurate for this model of human history.

\stdpopsim allows us to readily compare relative performance on this benchmark to
that based on a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but from the OutOfAfrica\_3G09
demographic model that does not include archaic admixture. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

To see how well methods might do at recovering the population history of a constant-sized population,
we can simulate under a human genome architecture and genetic map, but without demography.
We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

>>>>

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \renewcommand{\theHtable}{S\thetable}
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
        \renewcommand{\theHfigure}{S\thefigure}
     }
\newcommand{\stopsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{\arabic{table}}%
        \renewcommand{\theHtable}{X\thetable}
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{\arabic{figure}}%
        \renewcommand{\theHfigure}{X\thefigure}
     }

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but specifying such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and the possibility for error.
Population genetics, as a field, also lacks standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, \stdpopsim, that attempts to rectify this situation.
\Stdpopsim is a community-driven open source project, which provides easy access to a standard
catalog of published simulation models from a wide range of organisms
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature, community-based quality control pipeline.
We share some examples demonstrating how \stdpopsim
can be used to systematically compare demographic inference methods,
and we encourage an even broader
community of developers to contribute to this growing resource.
}
\date{}

\keywords{Population genetics, Simulation, Inference, Reproducibility}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past two decades. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{Boyko:2008cr,kim2017inference,tataru2017inference,Fortier703918,Huang2019,Vecchyo770966},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247,Barroso2019},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming ever more important.

Assessing the accuracy of inference methods for population genetics is
challenging in large part because the ``ground-truth'' in question
generally comes not from direct empirical observations, as the relevant
historical processes can rarely be observed, but instead from simulations.
Population genetic simulations are therefore critically important to the
field, yet there has been no systematic attempt to establish community
standards or best practices for executing them. Instead, the general modus
operandi to date has been for individual groups to validate their own
methods using bespoke simulations. 
Often these studies focus more on showcasing a novel method than on
rigorously comparing it with competing methods.
Moreover, this situation results in a great deal of duplicated effort,
and contributes to decreased reproducibility and transparency across the entire field.
It is also a barrier to entry to the field, because new researchers can
struggle with 
the many steps involved
in implementing a state-of-the-art population genetics simulation,
including identifying appropriate demographic models from the literature, 
translating them into input for a simulator, obtaining appropriate genetic
maps, and choosing appropriate values for key population genetic parameters.

A related issue is that it has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect population
genetic inferences. Yet there
are clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by other methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, some methods are likely to
perform better than others under certain scenarios, but 
researchers lack principled guidelines for selecting the best method for addressing 
their particular questions. The need for empirical
guidance will only increase as researchers 
seek to apply population genetic methods to a growing collection of non-model taxa.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains three organisms: humans,
\emph{Drosophila~melanogaster}, and \emph{Arabidopsis~thaliana}. For each
organism, the catalog contains details on the physical organization (e.g., chromosome structure)
of its genome, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.
In this way, \stdpopsim will lower the barrier to high-quality simulation for exploratory analyses,
enable rigorous evaluation of population genetic software,
and contribute to increased reliability of population genetic inferences.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/Figure1.pdf}
    \caption{\textbf{Structure of \stdpopsim}. \textbf{(A)} The
hierarchical organization of the \stdpopsim catalog contains all model simulation information
within individual species (expanded information shown here for \emph{H.~sapiens} only). 
Each species is associated with a representation of the physical genome, and one or more genetic maps and demographic models.
Dotted lines indicate that only a subset of these categories is shown. 
At right we show example code to specify
and simulate models using \textbf{(B)} the python API or \textbf{(C)} the command line interface.
    }
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library.}
The first contribution of the PopSim consortium is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig.~\ref{fig:cartoon}A), initially consisting of humans,
\textit{D.~melanogaster}, and \textit{A.~thaliana}.
A species definition consists of two key elements.  Firstly, the library defines
some basic information about each species' genome, including information about chromosome
lengths, average mutation rates, and generation times. We also provide access to detailed
empirical information such as genetic maps, which model observed
heterogeneity in recombination rate along chromosomes. As such maps are often large,
we do not distribute them directly with the software, but make them available
for download in a standard format. When a simulation using such a map is
requested by the user, \stdpopsim will transparently download the map
data into a local cache,
% (\texttt{$\sim$/.local/cache/stdpopsim} by default on Unix platforms)
where it can be quickly retrieved for subsequent simulations.
In the initial version of \stdpopsim we support
the HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps for humans;
the \cite{salome2011recombination} map for \emph{A.~thaliana};
and the \cite{comeron2012many} map for \emph{D.~melanogaster}.
Adding futher maps to the library is trivial.
The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which allow simulation under
specific historical scenarios that have been fit to present-day patterns of
genetic variation. (See the Methods for a description of the community
development and quality-control process for these models.)

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine.
We have implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim} 
as an alternative backend,
to allow simulation of processes that cannot be modeled under the coalescent.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work, and encourage authors
publishing demographic models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number and the long-term stable identifiers
for population models and genetic maps,
this citation information will result in well-documented and reproducible
simulation workflows. The individual tree sequence files produced by
\stdpopsim also contain complete provenance information including the command
line arguments, operating system environment and versions of key libraries
used.

@@

The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.
In addition to those models shown, at time or writing
the PopSim Consortium has models in development for \textit{Pongo abelii} and \textit{Escherichia coli}.

\renewcommand{\arraystretch}{1.2}
\begin{table}[t]
% Make sure the table is centered even though it's too wide
\makebox[\textwidth][c]{
    \begin{footnotesize}
    \input{model_table.tex}
    \end{footnotesize}
}
\caption{\label{tab:catalog}
Initial set of demographic models in the Catalog and simple benchmarks.
For each model we report the CPU time, maximum memory usage and the
size of the output \tskit file. In each case we simulate 100 samples
drawn from the first population, for the shortest chromosome of that species
and a constant chromosome-specific recombination rate.
The times reported are for a single run on an Intel i5-7600 CPU.
Computing resources required will vary widely depending on sample sizes, chromosome length,
recombination rates and other factors.
}
\end{table}

Currently, {\em Homo sapiens} has the largest number of population models in
\stdpopsim (see Table \ref{tab:catalog}).  
These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth; Africa\_1T12), the three-population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations (OutOfAfrica\_3G09), the \cite{tennessen2012evolution} two-population variant of the
\citeauthor{gutenkunst2009inferring} model which does not include Asian populations, but more explicitly models
recent rapid human population growth (OutOfAfrica\_2T12), the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components (AmericanAdmixture\_4B11), a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture (OutOfAfricaArchaicAdmixture\_5R19),
a complex model of ancient Eurasian admixture from \cite{kamm2019efficiently} (AncientEurasia\_9K19),
and a synthetic model of oscillating population size from \cite{schiffels2014inferring} (Zigzag\_1S14).
Together these models
contain features believed to have widespread impacts in real data (e.g., bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For other species we have have implemented two demographic histories for
\textit{D.~melanogaster}, three from \textit{A.~thaliana}, and have under development
model for \textit{P.~abelii} and \textit{E.~coli}.
For \textit{D.~melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample (African3Epoch\_1S16), as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring}, which jointly models African
and European populations (OutOfAfrica\_2L06). For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes (SouthMiddleAtlas\_1D17). We have also implemented a two-epoch and a three-epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene} (African2Epoch\_1H18 and African3Epoch\_1H18). 
%For \emph{P. pygmaeus} we have implemented the two
%species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
%and \emph{P. abelii}. 
%For \emph{E. coli} we have a simple, large constant population size model implemented.
In addition to organism-specific models, \stdpopsim also includes a generic piecewise constant size model and
isolation with migration (IM) model which can be used with any genome and genetic map.

To guarantee reproducibility,
we have standardized naming
conventions for species, genetic maps, and demographic models that will enable long term stability
of unique identifiers used throughout \stdpopsim,
as described in our documentation
(\url{https://stdpopsim.readthedocs.io/en/latest/development.html#naming-conventions}).

\subsection*{Use case: comparing methods of demographic inference}

As an example of the utility of \stdpopsim, we demonstrate how it can be used
to easily and fairly compare popular demographic inference methods.
Although we present comparison of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community
(see Discussion).

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To reproducibly evaluate and compare the performance of inference methods, we developed
workflows using \texttt{snakemake} \citep{koster2012snakemake}
that are available from \url{https://github.com/popsim-consortium/analysis},
that allow efficient computing in multicore or cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above.
Our workflow generates $R$ replicates of $C$ chromosomes, producing $n$ samples in each of a total of $R \times C$
simulations for each demographic model. After simulation, the workflow prepares
input files for each inference method by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairway plot}, and
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the OutOfAfricaArchaicAdmixture\_5R19 model and using the
HapMapII\_GRCh37 genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate.
As a proxy for the truth, in black we show inverse coalescence rates
as calculated from the true demographic model (see text).
}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

Figure \ref{fig:n_t_ragsdale} presents the results 
from simulations under OutOfAfricaArchaicAdmixture\_5R19,
a model of human migration out of Africa that includes archaic admixture 
\citep{ragsdale2019models}, along with an empirical genetic map. In each column of this figure
we show $N(t)$ inferred from samples taken from each of the three extant populations in the model.
In each row we show comparisons among the methods (including two sample sizes for \MSMC).
Blue lines show estimates from each of three replicate whole genome simulations.
There is no single ``true'' reference for effective population size
because of model misspecification---the inference methods are fitting a single population model
to data simulated from multiple populations.
However, many methods work by matching coalescence time distributions,
and a single-population model with varying population size can match any coalescence time distribution
(in which case coalescence rate is the inverse of the effective size).
For this reason, we used as our ``ground-truth'' (solid black lines) 
not historical census sizes,
but rather inverse coalescence rates calculated analytically in \texttt{msprime} (see Appendix).
While there is variation in accuracy among methods, populations, and individual replicates,
the methods are generally accurate for this model of human history.

\stdpopsim allows us to readily compare relative performance on this benchmark to
that based on a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but from the OutOfAfrica\_3G09
demographic model that does not include archaic admixture. Again we see that each
of the methods is capturing relevant parts of the population history, although the
accuracy varies across time. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can also
see how well methods might do at recovering the population history of a constant-sized population,
with human genome architecture and genetic map.
We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}.
All methods recover population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to infer sinusoidal
patterns of population size even though no such change is present.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/d_mel_Sheehan_mask2.png}
\caption{\textbf{Comparing estimates of $N(t)$ in \textit{Drosophila}}. Population
size over time ($N(t)$) estimated from an African population sample. Data were generated by simulating
replicate \textit{D.~melanogaster} genomes under the African3Epoch\_1S16 model
with the genetic map of \cite{comeron2012many}. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. 
As a proxy for the truth, in black we show inverse coalescence rates
as calculated from the true demographic model (see text).
}
\label{fig:n_t_sheehan}
\end{center}
\end{figure}

As most method development for population genetics has been focused on human
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from the African3Epoch\_1S16
model, originally estimated from an African sample of \textit{D.~melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \textit{D.~melanogaster} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting and generally worse than what we
observe under simulations of the human genome.
